Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-11-04 writedown/period.noneq.msm/paper.tex
adjust the plots

\documentclass[aps, pre, preprint,unsortedaddress,a4paper,onecolumn]{revtex4}
% \documentclass[aps, pre, preprint,unsortedaddress,a4paper,twocolumn]{revtex4}
% \documentclass[acs, jctcce, a4paper,preprint,unsortedaddress,onecolumn]{revtex4-1}
% \documentclass[aps,pre,twocolumn,unsortedaddress]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

Despite their high resolution in presenting a molecular system, MD the
trajectories including the time-dependent position and velocity of every atom
sometimes impose non-trivial challenge for data analyzing, in the sense that 
extracting the essential structure and dynamical properties is
becoming increasingly difficult as the  system size and trajectory length grow.
% by merely visulizing (not only molecular cartoons but also order parameters)
% and observing the molecules.
The Markov state model
(MSM)~\cite{A19-31}, which is a promising method for data analyzing of
the high-dimensional MD trajectories, has been well developed over the
past decade in theory and applications \cite{A19-31, A19-1}, and in
the software implementations~\cite{A19-49, MSMBuilder}.  The principal
idea of MSM is to approximate the original high-dimensional system by
a reduced Markovian dynamics over a finite number of states, from
which the essential structures and dynamics can be formulated by the spectrum analysis.
By assigning the spacial and temporal discretizations, 
the transition probabilities are sampled from MD trajectories, and
the MSM does not need more information from the MD simulation.
The quality of the MSM therefore relies on the quality of the discretizations.
Moreover, the characteristic of the original dynamics plays a role:
It has been shown that in principle good approximate accuracy would be achieved
by using only a few ``coresets'',
if the system is dominated by a few metastable states, in the sense that
the MD trajectories stay in the vicinity of a metastable state substantially longer than
transform to another such state~\cite{schuette2011markov}.
Besides the applications in analyzing the MD trajectories, it worth noting that MSM is
also used for designing initial configurations for massively parallel MD
simulations~\cite{kohlhoff2014cloud}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{fig/t010/discrete/fig-cg-prob.eps}  
  \caption{Time-dependent probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$.  The brute force NEMD simulation is compared with different
    spatial discretization methods. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp2}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{fig/t010/discrete/fig-cg-prob-tau.eps}  
  \caption{Time-dependent probability $\prob\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The NEMD simulation is compared with the Master equation using generators discretized with $N_\dih=20$ and different lag time $\tau$. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering  
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3.steadyDist//fig-dist.eps}
  % \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-dist-msm.eps}
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3/fig-floquet-vec-1.eps}
  \caption{The color scale plot of the logarithmic quasi-stationary distribution $\mymu$
    of (a) the NEMD  and (b)
    the discretized dynamics governed by Floquet transition matrix $\vect P=\myphi^{\top}(T)$.
  }
  \label{fig:num-1}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{fig/t010/cluster.marco.3/fig-cluster.eps}
  \caption{The coreset identification. Different colors indicate different core sets: $C_{\confaa}$ (green), $C_\beta$ (yellow) and $C_{\confc}$ (red).
    The blue color indicates the transition region $C$ that does not belong to one of the core sets.}
  \label{fig:cluster}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-1.eps}\\
  \vskip -.5cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-2.eps}\\
  \vskip -.5cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-3.eps}\\
  \caption{Comparisons of the first mean hitting time (FMHT) based on NEMD
    simulation (first column) and discretized dynamics (second column).  From top
    to bottom the first mean hitting times to the core sets $C_{\confaa}$, $C_{\beta}$,
    and $C_{\confc}$ are shown, respectively}
  \label{fig:num-6}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Forward and backward committors}
\label{sec:alanine-committor}
\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-1.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-1.eps}
  \caption{The forward $q^\fwd_{\confc}$ and backward committors
    $q^\bwd_{\confc}$ computed by the discretized dynamics (second row) and NEMD simulations (first
    row).}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-2.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-2.eps}
  \caption{The forward $q^\fwd_{\confaa}$ and backward
    $q^\bwd_{\confaa}$ committors computed by discretized dynamics (second row) and NEMD
    simulations (first row).}
  \label{fig:num-4}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-3.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-3.eps}
  \caption{The forward $q^\fwd_{\beta}$ and backward $q^\bwd_{\beta}$
    committors computed by discretized dynamics (second row) and  NEMD
    simulations (first row).}
  \label{fig:num-5}
\end{figure}

@@

\begin{table}
  \centering
  \caption{
    Comparison of second and third eigenvalues of $\vect P$ and $3\times 3$ MSM transition matrices $\hat{P}$ and $\hat{M}^{-1}\hat{T}$, respectively, from the two different MSM approaches.
  }
  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}c rrr}\hline\hline
      &  $\lambda_2$ & $\lambda_3$ & $\lambda_4$ \\\hline
    $P$                 &0.905  &0.668 & 0.551       \\
    $\hat P$     & 0.909  &0.710 & --       \\
    $\hat{M}^{-1}\hat{T}$    & 0.901  &0.649 & --       \\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp1}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{fig/t010/cluster.marco.3/fig-coreset-prob.eps}
  \caption{Comparison of NEMD and MSM long-term kinetics of alanine
    dipeptide in an oscillatory electric field. The plots show the
    time-dependent probability $\hat p_j(t)$ to be assigned to
    core set $C_j$ (corresponding to the observable $\mathcal A$ given
    in (\ref{A}) with $\alpha_j=1$ and $\alpha_k=0$ for $k\not=
    j$). Solid lines are from brute force NEMD simulations, while the
    dashed lines are from our MSM.}
  \label{fig:num-7}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{fig/flowchart/flowchart.eps}
  \caption{The flowchart of this work. A lighter color indicates
  a ``more coarse'' approximation to the original NEMD dynamics.}
  \label{fig:flowchart}
\end{figure}

>>>>

\documentclass[aps, pre, preprint,unsortedaddress,a4paper,onecolumn]{revtex4}
% \documentclass[aps, pre, reprint,unsortedaddress,a4paper,twocolumn]{revtex4}
% \documentclass[acs, jctcce, a4paper,preprint,unsortedaddress,onecolumn]{revtex4-1}
% \documentclass[aps,pre,twocolumn,unsortedaddress]{revtex4-1}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

Despite their high resolution in presenting a molecular system, MD the
trajectories including the time-dependent position and velocity of every atom
sometimes impose non-trivial challenge for data analyzing, in the sense that 
extracting the essential structure and dynamical properties is
becoming increasingly difficult as the  system size and trajectory length grow.
% by merely visulizing (not only molecular cartoons but also order parameters)
% and observing the molecules.
The Markov state model
(MSM)~\cite{A19-31}, which is a promising method for data analyzing of
the high-dimensional MD trajectories, has been well developed over the
past decade in theory and applications \cite{A19-31, A19-1}, and in
the software implementations~\cite{A19-49, MSMBuilder}.  The principal
idea of MSM is to approximate the original high-dimensional system by
a reduced Markovian dynamics over a finite number of states, from
which the essential structures and dynamics can be formulated by the spectrum analysis.
By assigning the spacial and temporal discretizations, 
the transition probabilities are sampled from MD trajectories, 
% the MSM does not need more information from the MD simulation.
so the quality of the MSM  relies on the quality of the discretizations.
Moreover, the characteristic of the original dynamics plays a role:
It has been shown that in principle good approximate accuracy would be achieved
by using only a few ``coresets'',
if the system is dominated by a few metastable states, in the sense that
the MD trajectories stay in the vicinity of a metastable state substantially longer than
transform to another such state~\cite{schuette2011markov}.
Besides the applications in analyzing the MD trajectories, it worth noting that MSM is
also used for designing initial configurations for massively parallel MD
simulations~\cite{kohlhoff2014cloud}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig/t010/discrete/fig-cg-prob.eps}  
  \caption{Time-dependent probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$.  The brute force NEMD simulation is compared with different
    spatial discretization methods. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp2}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig/t010/discrete/fig-cg-prob-tau.eps}  
  \caption{Time-dependent probability $\prob\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The NEMD simulation is compared with the Master equation using generators discretized with $N_\dih=20$ and different lag time $\tau$. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering  
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3.steadyDist//fig-dist.eps}\\
  % \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-dist-msm.eps}
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3/fig-floquet-vec-1.eps}
  \caption{The color-scale plot of the logarithmic quasi-stationary distribution $\mymu$
    of (a) the NEMD  and (b)
    the discretized dynamics governed by Floquet transition matrix $\vect P=\myphi^{\top}(T)$.
  }
  \label{fig:num-1}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3/fig-cluster.eps}
  \caption{The coreset identification. Different colors indicate different core sets: $C_{\confaa}$ (green), $C_\beta$ (yellow) and $C_{\confc}$ (red).
    The blue color indicates the transition region $C$ that does not belong to one of the core sets.}
  \label{fig:cluster}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3.fht/fig-fht-msm-3.eps}\\
  \vskip -.5cm
  \caption{Comparisons of the first mean hitting time (FMHT) based on NEMD
    simulation (left column) and discretized dynamics (right column).  From top
    to bottom the first mean hitting times to the core sets $C_{\confaa}$, $C_{\beta}$,
    and $C_{\confc}$ are shown, respectively}
  \label{fig:num-6}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Forward and backward committors}
\label{sec:alanine-committor}
\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-1.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-1.eps}
  \vskip -.5cm
  \caption{The forward $q^\fwd_{\confaa}$ and backward 
    $q^\bwd_{\confaa}$ committors
    and their difference $q^\fwd_{\confaa} - q^\bwd_{\confaa}$
    computed by the NEMD simulations (left
    column) and the  discretized dynamics (right column).}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-2.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-2.eps}\\
  \vskip -.5cm
  \caption{The forward $q^\fwd_{\beta}$ and backward
    $q^\bwd_{\beta}$ committors
    and their difference  $q^\fwd_{\beta} - q^\bwd_{\beta}$ 
    computed by the NEMD simulations (left
    column) and the  discretized dynamics (right column).}
  \label{fig:num-4}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-fw-msm-3.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-bw-msm-3.eps}\\
  \vskip -.7cm
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.3/fig-commitor-diff-msm-3.eps}\\
  \vskip -.5cm
  \caption{The forward $q^\fwd_{\confc}$ and backward $q^\bwd_{\confc}$
    committors
    and their difference  $q^\fwd_{\confc} - q^\bwd_{\confc}$ 
    computed by the NEMD simulations (left
    column) and the  discretized dynamics (right column).}
  \label{fig:num-5}
\end{figure}

@@

\begin{table}
  \centering
  \caption{
    Comparison of second and third eigenvalues of $\vect P$ and $3\times 3$ MSM transition matrices $\hat{P}$ and $\hat{M}^{-1}\hat{T}$, respectively, from the two different MSM approaches.
  }
  \begin{tabular*}{0.45\textwidth}{@{\extracolsep{\fill}}c rrr}\hline\hline
      &  $\lambda_2$ & $\lambda_3$ & $\lambda_4$ \\\hline
    $P$                 &0.905  &0.668 & 0.551       \\
    $\hat P$     & 0.909  &0.710 & --       \\
    $\hat{M}^{-1}\hat{T}$    & 0.901  &0.649 & --       \\
    \hline\hline
  \end{tabular*}
  \label{tab:tmp1}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.3/fig-coreset-prob.eps}
  \caption{Comparison of NEMD and MSM long-term kinetics of alanine
    dipeptide in an oscillatory electric field. The plots show the
    time-dependent probability $\hat p_j(t)$ to be assigned to
    core set $C_j$ (corresponding to the observable $\mathcal A$ given
    in (\ref{A}) with $\alpha_j=1$ and $\alpha_k=0$ for $k\not=
    j$). Solid lines are from brute force NEMD simulations, while the
    dashed lines are from our MSM.}
  \label{fig:num-7}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{fig/flowchart/flowchart.eps}
  \caption{The flowchart of this work. A lighter color indicates
  a ``coarser'' approximation to the original NEMD dynamics.}
  \label{fig:flowchart}
\end{figure}
